---
title: "Identification and Spatial Analysis of Russian Law Firms Naming Using Text Embeddings"
author: "Pavel Syomin"
output:
  word_document:
    reference_doc: "journal-article/reference.docx"
bibliography: "journal-article/refs.bib"
csl: "journal-article/apa.csl"
link-bibliography: False  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  echo = FALSE, 
  fig.width = 4.96, 
  fig.id = FALSE,
  dpi = 1200)
IS_PAPER <- TRUE # used in script to disable saving of names and samples
```

```{r package-imports, include=FALSE}
library(dplyr)
library(ggplot2)
library(glue)
library(here)
library(readr)
library(readxl)
library(sf)
library(stringr)
library(tidyr)
library(units)
```

```{r data-imports, cache=TRUE, include=FALSE}
source(here("journal-article", "Script.R"))
```

<div custom-style="Affiliation">

Perm State University, Geographical faculty, Department of social and economic geography, Perm, Russia

E-mail: ntsp\@ya.ru

ORCID: 0000-0002-4015-9206

</div>

<div custom-style="Abstract">

**Abstract:** The present study analyzes law firm naming in Russia, with a particular focus on the geographical aspects of this phenomenon. The primary objective of this research is to ascertain the impact of the spatial relations between the country's regions, specifically those of a neighboring, proximity, and belonging to a parent-level subdivision, on the naming of law firms. To conduct the analysis, data regarding law firms is extracted from the open dumps of the state registry of small and medium-sized businesses. The names of the firms are transformed into numerical vectors, subsequently clustered with the KMeans method, and then manually grouped to identify naming strategies. Numerical representations of law firm names within each region are averaged to get region embeddings, which are then used to compare the names of law firms in different regions. The research findings indicate that regions situated closer to one another or sharing a common border tend to adopt slightly more similar law firm names. Furthermore, regions that are part of the same economic region also display a greater degree of similarity in law firm naming. Three primary strategies of law firm naming are “Form”, “Law“, and “Service”. Volga and North Caucasus economic regions are characterized by the predominance of the “Service” strategy, whereas the Volga-Vyatka economic region exhibits a “Law” strategy, and the remaining territories demonstrate a combination of naming strategies. The results of the study lend support to the widely used classification of economic regions in Russia. The findings also underscore the impact of neighboring, proximity, and interconnectivity on the economic and cultural attributes of the regions. The administrative open data processing techniques, as well as the methods of working with firm names based on neural networks proposed in this paper, may be used by other economic geographers for their own research purposes.

</div>

<div custom-style="Keywords">

**Keywords:** firm names; law firms; text embeddings; economic regions, text similarity; large language models.

</div>

# Statements and Declarations

**Competing Interests.** No funding was received for conducting this study. The author has no financial or proprietary interests in any material discussed in this article.

\newpage

# 1. Introduction

The geography of services, also known as the geography of the service sector, is a crucial component of geographical studies in Russia. This field has a history spanning several decades, with origins dating back to the 1960s. The geography of services has two distinct dimensions: economic and social. The former is primarily concerned with the business aspects of the services, while the latter addresses the implications for human well-being arising from the geographical organization of services [@Tkachenko2016]. The existing scientific literature on the geography of services in Russia covers a broad range of topics, including education [@Shulgina2019], car rental [@Zyrianova2018], catering [@Preobrazhenskiy2018], medical care [@Kazantcev2022; @Panin2023; @Nikoloski2024], and internet access [@Nagirnaya2015; @Zemtsov2022]. However, some areas remain underexplored, with legal services being one such example [@Tkachenko2016; @Moiseeva2016]. 

Despite the lack of interest from geographers, the spatial aspects of the activity of law firms are worth investigating for a number of reasons. Firstly, legal services can be categorized as business or producer services, and this category of services is important for economic development, on both a national and regional scale [@Hansen1993; @Faulconbridge2019]. Secondly, despite their comparatively minor direct economic impact, these services play a significant role as intermediaries and facilitators, contributing to the establishment and maintenance of the rule of law, social order, stability, and predictability of social relations. Thirdly, for the general public, access to legal aid is vital for the protection of human rights and the attainment of social justice. In this regard, the spatial aspects of law services are on par with the spatial issues of medical care [@Blacksell1988; @Blacksell1990].

The study of firm names — which are sometimes also referred to as company names, brand names, or simply “brands” — is a relatively uncommon yet potentially fruitful approach to the geographical analysis of services and related problems. A firm name is designed to distinguish the business from competitors, to highlight the sector or geographical area in which it operates, and to convey information about the services offered by a provider and the associated benefits. A brand is therefore capable of transferring a substantial amount of information in a brief textual format. Such information may prove useful in uncovering the city's business landscape and consumer perceptions of a particular city with its distinctive services [@Raento2001]. It may also help to analyze trends common to a specific sector in a geographic region [@Marjavaara2022]. The incorporation of place names (toponyms) in brands influences the business development and consumer attitudes [@Peluso2012; @Zhang2014; @Erdogan2016; @Noorda2019]. Company names encapsulate a multitude of intertwined properties relevant to the economy and culture of a given location. This synthesis makes them suitable for the identification of regions, especially in the case of vernacular regions [@Reed1976; @Liesch2015; @Vukosav2016; @Holtkamp2018].

It seems reasonable to posit that (1) several law firms naming strategies may be identified in Russia and (2) there is a spatial heterogeneity in the distribution of the firms with different naming strategies. Law firms naming strategy here is defined as a typical model, or pattern, of brand construction identified by usage of identical or semantically similar words or phrases. The potential viability of these two statements cannot be directly conducted from the existing research, but it relies on a common-sense observations about the similarities and differences of law firm names in and between various regions of the country. Law firm names are likely to reflect some economic and, to a lesser extent, cultural characteristics of the territories where these firms are located. Thus, given the vast area and a material heterogeneity of economical and cultural landscape of Russia, these names should not have a uniform spatial distribution either. The primary objective of this paper is to discover these hypothesized naming strategies and their spatial heterogeneity (if any). In addition, it aims to test whether the variation of law firms naming between settlements in Russia is contingent upon the geographical proximity of settlements and economic links between them. Building upon the Tobler's First Law of Geography (“near things are more related than distant things”) [@Tobler1970], it can be hypothesized that both these factors have a positive impact on the similarity of law firm names, because small geographic distance directly makes settlements “near”, as well as economic links make them to be “near” in economical meaning. Finally, due to quite a pilot nature of this paper, it also tries to handle all intriguing side observations about the law firms naming strategies and their spatial distribution in Russia that may arise from the exploratory analysis of the data, even if no robust testing can be performed.

# 2. Materials and methods

## 2.1. Data

The issue of obtaining information about firm names is a challenging one. All Russian organizations are registered in the state registry operated by the Federal Tax Service (FTS of Russia). An extract from this registry about a particular company can be searched and downloaded online by any person free of charge. However, a considerable sum of money must be paid and a written request submitted to the FTS of Russia to purchase a complete copy of the registry. Commercial companies offer data resell services, but the prices are similarly high.

To address these challenges, a unified state registry of small and medium-sized enterprises (SME registry) may be used [@SMERegistry2024]. The registry has been maintained since 2016 in accordance with Article 4.1 of the Federal Law of 24 July 2007 No. 209-FZ “On the development of small and medium-sized businesses in the Russian Federation”. As a rule of thumb, all companies and individual entrepreneurs with a yearly revenue below 2 billion roubles and an average yearly number of employees below 250 are automatically included in the SMB registry, without any action being required on their part. There are additional qualifications for the SME registry, e. g. absence of non-SME or foreign control such as, but these should not affect a material portion of law firms. The registry is fully downloadable at the FTS open data portal at no cost. Additionally, the portal provides a monthly record of updates made since the registry's inception. To sum it up, the SMB registry can be regarded as the most comprehensive free data source offering information on Russian business entities, both organizations and individuals.

The downloadable version of the registry includes, in particular, firm names, addresses of registration (up to a settlement), and codes of activity kinds compatible with NACE Rev.2. It should be noted that the registry does not include several types of firms, e.g. foreign commercial law firms without a juridical person registered according to Russian law, non-profit organizations, bar (advocacy) establishments, large companies etc. This results in some data being absent. Nevertheless, this issue should be considered insignificant for a variety of reasons. Foreign law firms usually register their affiliated juridical person in Russia and thus are not omitted in the registry. Non-profit organisations' share on Russian legal market does not exceed 1% [@Moiseeva2016]. The names of bar establishments are usually quite uniform, containing the kind of the establishment (cabinet of an advocate, bureau of advocates, college of advocates, juridical consultancy) accompanied with a personal name(s) of advocates or with a mention of geographical location of a particular establishment, thus their potential spatial heterogeneity is not characteristic. A matching of the list of the law firms obtained from the SME registry with the extract of law firms from the Russian Financial Statements Database shows that ~3,300 of the firms present in SME registry in 2023 are absent from RFSD, and ~5,200 firms present in RFSD are missing in SME registry. The balance of missing records is not that far from the equality, and with the total number of law firms in SME registry (~34000) the share of missing entries is below 15%. Consequently, the completeness of information is relatively high, and it can be used to draw reliable conclusions.

## 2.2. Methods

### 2.2.1. Law firms selection criteria

In this study, the term “law firm” is defined as a commercial organization registered in Russia in accordance with the Civil Code of Russia, with a NACE Rev.2-compatible main activity code equal to 69.10 (activity in the area of law). The aforementioned definition excludes individual entrepreneurs, foreign companies, non-profit organizations, advocacy firms, and notaries. The selection technique based on the main activity code is derived from that proposed by @Moiseeva2016. Individual entrepreneurs are excluded from the analysis because they typically operate under their personal names, and even if they have a brand, it is not listed in the SME registry. Other types of law firms are not included in the analysis primarily due to data availability issues.

The data used in the study covers a period of six years, from 2016 to 2021. The initial point in time is defined by the commencement of SME registry data availability. The finish point, 2021, is the latest year, is the latest year with “normal” conditions in Russia not impacted by military operation in Ukraine and consequent international sanctions and domestic economic and political perturbations. The purpose of temporal dimension is to exclude short-living law firms that are likely to be dummy companies. Organizations with a lifespan of less than three years have been omitted from the research, while all the remaining have been included in the further analysis.

### 2.2.2. Firm data acquisition

The processing of SME registry data is conducted in a series of stages. In the initial phase of the research process, all source archives are downloaded. Each source archive contains a full copy of the registry data for the specified publication date, with a new archive becoming available on a monthly basis. Subsequently, data pertaining to law firms is extracted from the archives and stored in CSV files. Each CSV file contains data from one source archive. Subsequently, all CSV files are consolidated into a single table, the duplicated records are removed, and the remaining records are augmented with start and end temporal markers. Subsequently, the addresses are normalized to resolve the inconsistencies in location names and geocoded up to a city/settlement level to add geographical coordinates. The lack of within-city location in the source data prevents from more fine-grained geocoding but does not pose a serious problem because city-level coordinates are enough for large-scale analysis on country or region level. Finally, the resulting table is transformed into a panel view, with each row representing a single law firm at a given year. All these stages are implemented in an open-source Python command-line tool.

### 2.2.3. Firm name embeddings retrieval

The Yandex Foundation Models cloud service [@Yandex2024] is used to transform the firm names into text vectors, or embeddings. The “text-search-query” model is selected because its purpose, as stated in the documentation, is to vectorize short texts. Prior to vectorization, the firm names are preprocessed to remove indications of organizational form such as “limited liability company”. This is done because such indications are common to the majority of names and thus have little value. The results of the queries to the Yandex API are stored separately for reuse, given that API calls are paid and take time to be completed. The dimension of a firm name vector is 256. Contextual embeddings (YandexGPT) were used in favor of static ones (e.g. FastText), because during the preliminary analysis, they have resulted in far better clusterisation of law firm names (see the description of this step below) which is crucial for naming strategies identification.

### 2.2.4. Naming strategies identification

The dimension of law firm names vectors is reduced from 256 to 2 with the UMAP method, and the resulting 2D vectors are clustered using the DBSCAN technique and plotted to illustrate the results. The clusters are subsequently subjected to manual analysis. A random sample of 20 names is selected from each cluster. The names samples are carefully examined, and each cluster is attached a concise description (e.g. “Consulting companies”), a subcategory, and a category. The mapping of clusters and their respective subcategories and categories (e.g. cluster N — “Legal bureau” / “Lawyers”) is employed to link all the firm names to the appropriate groups. The concept of naming strategies is derived from the works of @Arora2015 and @BassolsiGardella2023. However, the specific groups differ from those used in their respective papers and are drawn from the data-driven analysis. The specific descriptions, subcategories and categories are provided, explained and analysed in the results section below.

### 2.2.5. Spatial regularities identification

Analysis of autocorrelation is used to check whether the spatial distribution of names belonging to each cluster identified on the previous step is random or has some regularities. Regions (constituent entries) or the Russian Federation were selected as a spatial subdivision for calculating the autocorrelation measures. It could have been more reasonable to use the next spatial level, municipal territories, to perform the computations, but exploratory data analysis has shown that a lion's share of such territories has no law firms, because the legal business is highly concentrated in cities and large settlements. 

The autocorrelation analysis has two branches differing in the way the spatial weights matrix is constructed. In the first branch, neighborhood of the regions is defined based on the adjacency of region's polygons. Sakhalin and Kaliningrad regions, being enclaves and thus having zero immediate neighbors, were manually connected with a few closest regions. In the second branch, the neighborhood of the regions was identified based on their belonging to the same economic region according to the classification widely used by economic geographers and spatial development practitioners in Russia for decades and outlined, for example, in [@Leyzerovich2010]. All regions within the same economic region are considered to be connected, and all regions of the different economic regions are disconnected, resulting in spatial weights graph comprised of several independent components. Despite the fact that the concept of economic regions in Russia is sometimes considered legacy, it is still implicitly used by researchers and practitioners, and the belonging to the same economic regions implies economic linkage, similarity and cooperation of regions. Thus, it makes sense to utilize this classification.

For each cluster, a global Moran's I metric is calculated. The variable of interest is a share of law firm names belonging to the cluster in a total quantity of law firm names in a region. If there are no law firm names of the cluster in the region, the share is considered to equal zero. P-values are adjusted for multiple comparisons with Holm's correction. Significance level (alpha) is set to 0.05.

For the clusters with statistically significant global Moran's I, a local Moran's I was subsequently calculated. P-values are adjusted with local Holm's correction. Cluster classification of local Moran's I are mapped.

# 3. Results and discussion

## 3.1. Law firms naming strategies

Law firm names clusters resulted from UMAP → DBSCAN pipeline are illustrated using scatterplot in Fig. 1. The visual quality of clusterisation is high: all the clusters are compact, round, and separated by considerable gaps. There are no noise points, and all the names have been easily assigned to clusters. 

Table 1 shows cluster sizes and sample names from the clusters with manually assigned short cluster descriptions, categories and strategies. The quantity of clusters (52) is relatively big, and their size is not even: while the largest group of names consists of items, the smallest one includes just 22 objects. A substantial number of clusters (...) is comprised of a mixture of short (a single word) names that cannot be given a reasonable description. Their strategy is marked as “Misc” and excluded from further analysis. “Misc” strategy encompasses ... law firm names. All the other names are split between three primary strategies: “Helpers”, “Lawyers”, “Service providers”. “Helpers” are law firms that articulate aid, protection, guarantee, advise, agency in their brands. Such names may either directly contain words related to legal aid or protection (“Aid” and “Protection” categories) or indirectly cause safety-related connotations by using the words associated with professionals or communities that do not leave a person alone (“Guarantor”, “Expert”, “Agent”, “Partner”, “Association” categories). In contrast to “Helpers”, “Lawyers” make a strong emphasis on their belonging to the legal services market by using the words “law”, “lawyer” or various Russian equivalents of “law firm” phrase. A distinct category of law firms with this strategy is formed by organisations that use a personal branding that follow the broad pattern “(law firm) <last name of a person> with partners” where (law firm) is an optional part, and <last name of a person> is a pointer to a featured lawyer of this organisation. Even if “law firm” subpart of the firm name is optional, the entire pattern is strongly associated in Russia with lawyers, thus this group of names is assigned the “Lawyers” strategy. Finally, “Services” strategy encompasses companies with a mention of a target audience (e.g. “Business”) or industry (e.g. “Construction”), a particular kind of services (e.g. “Consulting”, “Expertise”, “Audit” or “Finance&Investment”) or services in general. Compared to two other primary strategies, this one usually do not directly mentions law and legal services market. It is likely that some of companies following this strategy, in fact, do not provide legal services, focusing instead on paralegal activity or using standard law firms NACE code by a mistake or as a false flag. It should be noted that in some cases, a category of names within the “Services” strategy point to a law firm specialization (e.g. consulting or real estate), so the further analysis of names may be used to distinguish between full service and specialized law firms and between particular specializations.

## 3.1. Law firm names in neighboring regions are more similar 

The semantic distance between neighboring regions in comparison to the semantic distance between regions that do not have a common border is illustrated in Fig. 1. Although the difference in medians is relatively minor, it is nevertheless present, and the overall distribution of distances between non-neighboring regions is shifted towards the area of higher values. Therefore, neighboring regions have more similar names of law firms than regions that do not border one another.

```{r fig-1-neighbors, fig.cap="**Fig 1.** Semantic distances between neighboring and non-neighboring regions", fig.height=3}
neighbors_plot
```

## 3.2. Greater geographical distance reduces the similarity of law firm names

The relationship between geographical and semantic distances is shown in Fig. 2. The correlation coefficient is relatively low but positive. Thus, regions that are situated close to each other tend to have a bit more similar names of law firms even in the absence of a common border between them.

```{r fig-2-distance-plot, fig.cap="**Fig. 2.** Relationship between geographical and semantic distances between regions", fig.height=3}
distance_plot
```

## 3.3. Law firm names inside an economic region are more homogeneous

Within-distances and outside-distances by economic regions, as outlined in the methods section, are depicted in Fig. 3. In some instances, the semantic distance between regions that are part of the same economic region is slightly less than the semantic distance between these regions and regions of other economic regions. However, this pattern does not apply to at least three economic regions: in Volga, North, and North Caucasus, median within-distance is equal to or greater than median outside-distance. Consequently, the results of this test are mixed: while links between regions constituting an economic region usually slightly increase the similarity of law firm names, there are several exceptions to this rule. A greater within-distance compared to outside-distance may indicate internal heterogeneity within an economic region, thereby suggesting the need for a classification update.

```{r fig-3-economic-regions-distance, fig.cap="**Fig. 3.** Semantic distances between regions within and outside economic regions", fig.height=5}
economic_regions_distances
```

## 3.4. Geographical patterns of naming strategies are unclear

The map in Fig. 4 depicts a dominant naming strategy in regions of Russia. As “Form” naming is used by a minority of law firms across all regions, it is not shown. European South of Russia and Povolzhye are characterized by a prevalence of the “Service” strategy. In other parts of the country, an irregular mixture of both “Law” and “Service” naming strategies is observed.

```{r fig-4-naming-strategies, fig.cap="**Fig. 4.** Map of naming strategies"}
naming_strategies
```

## 3.5. Naming strategies within economic regions are inconsistent

As seen from Fig. 4, few economic regions are characterized by a uniform naming strategy. In the Volga-Vyatka economic region, the “Law” strategy is the most prevalent among all the regions, with the exception of the Republic of Mari El. In Volga and North Caucasus economic regions, service-based strategy is the prevailing approach in the majority of regions. The remaining economic regions exhibit a combination of law firms naming strategies. 

# 4. Conclusion

The findings of the study are somewhat contradictory. On the one hand, several hypotheses have been proven. These include the following: law firm names in neighboring regions are more similar than law firm names in non-neighboring regions; the smaller the geographic distance between regions, the more similar the law firm names; and law firm names within an economic region are typically more similar compared to the law firm names outside an economic region. On the other hand, the difference in the metrics of similarity is relatively small in each case, as is the correlation coefficient. It is therefore important to note that although the postulated phenomena do indeed exist, their magnitude is weak.

The three principal strategies for naming law firms are as follows: “Form”, “Law”, and “Service”. The first of these represents a minority approach, while the remaining majority is divided between two other strategies. In the European South of Russia and in the Volga economic region, “Service” is a dominant strategy. However, no other spatial patterns in law firms naming have been identified. Economic regions are typically characterized by a mixture of naming strategies, except for Volga-Vyatka, where the “Law” strategy prevails, and Volga and North Caucasus, where the “Service” strategy is dominant.

The contribution of this study is primarily theoretical and methodological in nature. It underscores the significance of firm names as a source of information for business- and economy-related research. It also proposes a technique for comprehensive analysis of branding in a geographical context through the use of modern natural language processing (NLP) solutions based on artificial neural networks. Additionally, it advocates for the utilization of disaggregated administrative open data for a spatial analysis. The findings demonstrate the effect of geographical proximity, neighboring, and economic relations on the similarity of the economic and, to a certain extent, cultural properties of the territories represented in firm names. The results generally align with the widely accepted classification of economic regions in Russia, although they highlight a few instances where a theoretically undesired internal heterogeneity within an economic region is observed. Finally, the study provides an illustration of a semi-automated brand classification within a specific economic sector and discovers a geographical variation in naming strategies that could be further investigated.

# Data availability

A dataset about Russian law firms, vectors of law firm names obtained from Yandex Foundation models, results of manual labelling, additional files used for mapping, as well as all the code used to conduct the study and generate the paper file are available at Zenodo: https://doi.org/10.5281/zenodo.13691418.

# References

<div id="refs"></div>
