---
output: 
  word_document:
    reference_doc: reference.docx
bibliography: refs.bib
#csl: style.csl
link-bibliography: False
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.id = FALSE,
  fig.width = 6.7,
  fig.height = 5,
  dpi = 300)
```

```{r imports, include=FALSE, cache=TRUE}
source("Script.R")
```


**Pavel O. Syomin[^1]**

<br>

<div custom-style="Title">

Legal Business Geography in Russia: Regional Analysis Using Tax Service Open Data

</div>

<br>

# Abstract

This paper provides an exploratory regional analysis of legal business in Russia for 2016–21. The analysis is based mainly on the Federal Tax Service (FTS) open data on small and medium-sized businesses and on the number of employees of organisations, as well as on portions of official statistical information. The paper proposes a reproducible technique of working with FTS open data for regional analysis in service geography that was partially implemented in a specially designed Python command-line tool. The analysis includes questions about the count of legal firms and their employees by region, about the share of organisations in regional legal business, about the role of legal business in region's economy, and about the yearly change of the number of legal companies in regions. In addition, it briefly addresses an issue of probable connection between the location of multi-regional courts and the development of legal business. It is found out that legal business in Russia is highly concentrated in Moscow and Saint Petersburg that jointly account for 1/3 of legal firms and workers and are also at the top by the role of legal business in the region. However, there are other regions with relatively high development of legal business, that are located mainly on the European South of Russia, in the Volga-Urals and in Siberia. There is a group of regions with domination of organisations in legal business and the group of regions where individual entrepreneurs prevail over companies. District courts are usually located in the regions with high level of legal business development, but the mechanism of this co-location requires further investigation. In general, the number of legal firms in regions grown from 2016 to 2021, with the common inverted U-shape pattern probably explained by the coronavirus (covid-19) pandemic that negetively impacted the small and medium business and reverted the initial growth trend.

<br>

<div custom-style="Block Text">

**KEYWORDS:** regional analysis, geography of services, legal services, legal business, open data

</div>

<br>

[^1]: Perm State University, Department of Social and Economic Geography; 15 Bukireva st., Perm, 614068, Russia; e-mail: ntsp\@ya.ru

# Introduction

Geography of services is an important branch of modern economic and social geography. In Russia, it dates back to the 1960s when the pioneering research papers in this field of studies were published [@Tkachenko2016]. In foreign countries, the spatial aspects of services were treated as a part of human geography and touched in the scholarly works about retail [@Reilly1931; @Berry1967] and cities [@Brush1953] usually based on the central places theory [@Christaller1966]. A series of review research articles published in 1985–87 directly uses the term “geography of services” and cites dozens of original papers related to this topic [@Daniels1985; @Daniels1986; @Daniels1987]. Newer foreign studies in this field focus on the processes and phenomena in peripheral or smaller economies [@Hermelin2007] or on the arising and growing sectors like FinTech [@Lai2020; @Wojcik2020]. In Russia, recent research in the geography of services covers spatial properties of banking [@Shustova2015], education [@Shulgina2019], car rental [@Zyrianova2018], food service [@Preobrazhenskiy2018], service sector in general [@Lisin2020]. However, several types of services are rarely studied, e. g. business and legal ones [@Tkachenko2016].

The research on the geography of legal services in Russia is limited. To date, only an analytical review describing the legal market in Russia was published that contains a chapter on its spatial aspects [@Moiseeva2016], and yet another paper briefly addresses the role of Russia in the global legal services market without diving into the internal spatial properties of the Russian legal services itself [@Yastrebov2021]. There are also non-scientific products about Russian legal market that include its regional analysis[^2]. Despite the observed low interest to this issue from other researches, it can be argued that the geographical studies of legal services is necessary. First, legal services, as a subset of other business or producer services, play a substantial role in the development of regions and territories and in the economic growth [@Hansen1993], and sometimes act as a limiting factor, that is, the lack of professional legal advice may cause problems to entrepreneurs, especially to small ones, thus decreasing their performance [@Hitchens1997]. Next, the in-depth analysis of legal companies may give valuable insights about the business sectors they serve [@Faulconbridge2019] or increase the theoretical understanding of markets and industries in general [@Henderson2016]. Last but not least, legal services, apart from their narrow economic function, should be also treated as one of the elements of legal system in a wider sense that protect human rights and freedoms, ensure the rule of law, and act as one of the intermediaries between people, business and government that help to resolve the conflicts, increase the certainty and orderliness of social relations. This aspect of legal services is closely related to the social dimension of geography of services, and from this point of view it is not surprising that the research is also being conducted on the problem of legal services availablity [@Baxter2015].

[^2]: Анализ рынка юридических услуг в России в 2019-2023 гг, прогноз на 2024-2028 гг. Web resource: https://marketing.rbc.ru/research/41683/ (accessed 14.04.2024)

Legal services is quite a narrow and specific area of economy that does not reported separately in the official statistics, so for their analysis, other sources of information are necessary. Company-level georeferenced data is valuable for this task, because it can be easily filtered to leave the firms doing the business in the required activity area only and then aggregated by its spatial attributes to make a geographical overview. Researchers typically extract such the information from public GIS services [@Lachininsky2021] or business intelligence (BI) systems [@Moiseeva2016]. The disadvantage of this method, however, is the complexity of data preprocessing or the relatively high price of the data. Also, GIS and BI systems often do not enable access to historical data, providing real-time picture only, while the geographical studies may benefit from combining spatial and temporal properties. To overcome all these issues, an open administrative data may be used. In particular, the publicly available archives of Registry of Small and Medium-Sized Businesses (SMB registry) distributed by the Federal Tax Service (FTS) of Russia, in combination with a few other open data collections of the same governmental body, look valuable for the task in hand, and they have already been used by geographers [@Rostislav2021; @Gumenyuk2022]. 

The purpose of this paper is twofold. On the one hand, it aims to provide an overview of geography of legal business in Russia. The scope of the research is intentionally limited due to various organizational and technical constraints: it deals with the service providers only (the supply part of the market) and contains nothing about customers (the demand part), focuses on commercial legal companies with other market participants such as notaries, advocacy firms, non-profit legal organizations left untouched, and operates with only a portion of variables and metrics indicating the characteristics of legal business. On the other hand, the paper proposes a reproducible technique of working with FTS open data for analyzing the geography of services in Russia, particularly the services that do not reported independently from the others in the official statistics (mainly specific types of business and customer services) and thus can hardly be analyzed with the traditional economic geographers' toolkit.

<br>

# Materials and methods

**Data.** The research utilizes two primary data sources. All of them are published on the FTS website as open data.

The first dataset is called *“Registry of small and medium-sized enterprises”*[^3]. It contains full dumps of the respective registry in zipped XML files. The first dump was published in August 2016, and since then, new dumps were published every month. Each dump contains the whole registry and includes, in particular, company's names, their taxpayer IDs, activity codes according to All-Russia Classifier of Kinds of Economic Activity, address of incorporation (up to a settlement name). According to Russian laws, the Registry includes information about all commercial organizations and individual entrepreneurs that meet several criteria, the most important of which are limit on revenue (no more than 2 billion Russian rubles) and average number of employees (no more than 250). Thus, the open dumps of this registry can be treated as the most complete publicly available data source on the Russian business entities. In addition, this source has geographical reference in the form of human-readable addresses, and historical records from 2016 to now in the form of time series of monthly dumps of the registry. 

[^3]: Единый реестр субъектов малого и среднего предпринимательства. Web resource: https://www.nalog.gov.ru/opendata/7707329152-rsmp/ (accessed 15.04.2024)

The second dataset is called *“Information about the average list count of employees of organisations”*[^4]. As the name suggests, it contains information about the number of company's employees. The data was initially published in 2019 and since then was updated with varied frequency, but at least yearly. As a result, the data archive includes information for the years from 2019 to 2023. For the unspecified reasons, the old versions of the dataset containing information for years from 2019 to 2021 were once deleted from the FTS website, but they had been already downloaded by the author and thus used in this research.

[^4]: Сведения о среднесписочной численности работников организации. Web resource: https://www.nalog.gov.ru/opendata/7707329152-sshr2019/ (accessed 15.04.2024)

In addition to the primary data sources, two lookup tables with the information on Russian cities and settlements are used to normalise and geocode raw addresses contained in the FTS data. The first table is called *“Settlement of Russia: population and geographical coordinates”*[^5]. It was prepared by the non-profit open data initiative called “Research data infrastructure” (RDI). The second table called *“Cities of Russia”*[^6] was published by commercial company DaData. Both these tables contain information that is actual for the beginning of 2022, thus there may be some inaccuracies in geocoding and normalisation, but they are likely to be negligible.

[^5]: Населенные пункты России: численность населения и географические координаты. Web resource: https://data.rcsi.science/data-catalog/datasets/160/ (accessed 15.04.2024)
[^6]: Города России. Web resource: https://github.com/hflabs/city (accessed 15.04.2024)

**Methods.** A special Python command-line tool called *ru-smb-companies* was developed by the author of this paper to initially process huge and complex raw datasets from the FTS servers and transform them into a single table. The source code of this tool with a brief documentation is published in author's personal GitHub account[^7]. Here, a concise summary of the data processing workflow is provided. The data processing takes several steps.

[^7]: Russian small and medium business dataset generator. Web resource: https://github.com/PavelSyomin/russian-smb-companies (accessed 15.04.2024)

*First,* source datasets were downloaded from the FTS servers and stored on the local machine. In addition, their backup was uploaded to author's personal Yandex.Disk account, making it possible for other researchers to access the raw data even if FTS decides to remove it (as indeed happened with data on the number of employees).

*Second,* the filter-on-extract approach was used to extract necessary information from the raw archives into intermediary CSV tables. The filter-on-extract approach means that instead of unpacking the entire archive on the disk, the program reads its contents file by file without unpacking, parses the internal XML structure, extracts necessary attributes only, filters the records by activity code equal to 69.10 to leave legal companies only, and stores the result on the disk. Thus, the runtime and disk usage are substantially reduced compared to the traditional extract-then-filter method. 

*Third,* the data from the intermediary CSV tables was aggregated to remove duplicated records. The reason for the duplicates is the procedure of data publication: as it was mentioned earlier, every version of the source dataset contains the full data for the time of publication rather than updates only, leading to the huge number of repeating records.

*Forth,* the addresses were normalised and geocoded to convert original administrative address format containing region, district, city and settlement name into simpler yet useful triplet of region, area and settlement, and add geographical coordinates and municipal codes (ОКТМО in Russian) for the settlements. 

*Fifth,* the data on business entities was combined with the data on employees using taxpayer ID as unique identifier common to both datasets, and also converted into a panel time series table that includes yearly information for each company or individual. This resulting table is used in the further analysis.

The research follows the general framework of regional analysis and uses the methods of descriptive statistics and data visualization, including the plotting of data on a tile-grid map of Russia. The analysis consists of a few research problems, or questions, that are listed below with the description of the relevant techniques used to find answers to them. 

The first problem is the *distribution of legal firms by regions* of Russia. To answer this question, legal firms that existed in 2021 are counted by the region, and the result is visualized on the map.

The second problem is the *structure of legal business in regions* with respect to the kind of firms (either individual entrepreneurs or organisations). To answer this question, legal firms are counted by region and king, and the result is visualized on a geofacetted plot.

The third problem is the *role of legal business on the regional and national scale*, that is, either its importance within a given region or the importance of region's legal services in the Russia's legal market. To discover this issue, two calculations are made. The first one is used to better understand the role of the region in the country's legal business and is done by calculating the share of legal companies employees of a region in the total number of country's legal companies employees. The second one is aimed at estimating the relative role of legal business in region's economy and is done by dividing the number of legal companies employees in a region by the total number of employees in a region. Both two metrics are then plotted on a tile-grid map.

The fourth issue deals with the temporal dimension and reflects the *change of the number of legal firms* in the regions. Legal firms are calculated by region and year, and the results are plotted.

Several aspects are common to all the questions. First of all, despite the presence of data from 2016 to 2024, only a part from 2016 to 2021 inclusively is utilized for the analysis. The reason for this is the Russia's military operation in Ukraine and the subsequent structural transformation of the economy, international sanctions, substantial political and legal changes in the country that may have significantly impacted the legal market compared to the previous period, and even though all these changes could have been interesting to study, the purpose of the current research is to draw an overall picture of legal market in Russia in “normal situation” without any serious perturbations. Statistics for several consequent years, if not processed separately (e. g. in the case of temporal changes), is averaged using median as a function (e.g. number of legal firms for 2016–2021 in the region is the median of counts for each year from 2016 to 2021). Next, both organisations and individual entrepreneurs are accounted for. If statistical computations require the count of employees, it is assumed that an individual entrepreneur always has exactly one employee, that is, this person itself. This is because even though individual entrepreneurs in Russia can legally employ people, the corresponding information is not available in the FTS open data possibly due to personal data protection issues. Also, if an organisation has reported zero count of employees, it is still assumed that is has exactly one employee, that is, the company's owner itself. 

<br>

# Results and discussion

**Average yearly count of legal firms by region** is displayed on a tile-grid colorpleth map in Fig. 1. Several conclusions can be drawn from this visualisation. First, Moscow is the only city where the number of legal firms is above 10 thousand, and this is the highest value in Russia. Next, a set of regions also has relatively high count of legal companies and individual entrepreneurs. In spatial terms, most of these regions form several groups. One of these groups is located at the south of the European part of Russia and consists of Krasnodar and Stavropol territories with Rostov, Volgograd and Voronezh regions. Tatarstan and Bashkir republics with Samara, Sverdlovsk, Chelyabinsk regions and Perm territory make up another group that may be called Ural-Volga. Some Siberian regions, namely Tyumen, Omsk, Novosibirsk, Kemerovo, with Krasnoyarsk territory, form the third group that may be named Siberian by its location. Also, Saint Petersburg city, Nizhniy Novgorod region, and Primorie territory act as single-region groups. The majority of other regions has relatively low count of legal firms, and in some of them (several republics of North Caucas and Altai and a portion of regions from the North) there are less than a hundred of such companies. The co-loocation of regions and groups with different count of legal firms suggests that there may be a sort of central-peripheries relations between them. For example, Moscow city and region, Saint Petersburg city, and Nizhniy Novgorod region are surrounded by dozens of regions with relatively small number of legal firms, and thus may act as a centers of legal business for them. The same may be true for the pairs of South and Caucasian, as well as for Siberia and Far East. It may be hypothesized that economic development of regions is positively associated with the development of legal business. Also, the population of a region may also matter.

<br>

```{r fig-1-counts-map, fig.cap="Fig. 1. Median yearly count of legal firms (both organisations and individual entrepreneurs) by regions of Russia for 2016–21. Grey is no data"}
counts_map
```

<br>

Fig. 2 displays the *median shares of organisations among all legal firms in regions*. Legal firms may be either organisations or individual entrepreneurs, so for each region, the rest part of the firms consists of individuals. The share of companies among all legal firms varies from less about 25% to almost 70%. However, a substantial part of regions has from 40% to 60% of organisations, and thus many regions may be categorized either as with a small domination of companies or as with small domination of individuals on the legal market. Several southern regions (Krasnodar and Stavropol territories, Rostov region, Kalmyk republic), as well as several northern ones (Arkhangelsk region, Komi republic, Nenetz and Yamal territories), and some central regions (Kaluga and Moscow) have substantially lower shares of companies compared to the surrounding regions, thus forming the areas of predominantly individual legal business. The overall count of legal firms seems to have to connection with the share of organisation among them.

<br>

```{r fig-2-shares-map, fig.cap="Fig. 2. Median yearly shares of companies among legal firms by regions of Russia for 2016–21. Grey is no data"}
shares_map
```

<br>

Fig. 3 shows the joint results of regional analysis of number of legal companies employees. Two main metrics are drawn, the first is the concentration of legal workers in the region, that is, the share of legal companies employees of a region in all country's employees of legal companies, and the second is the share of legal workers in region's labor force. The first measure roughly describes the importance of a region in country's legal business, while the second estimates the legal business' impact in regional economy. In addition, the number of multi-region courts located in a given region is shown. The overall regional distribution of legal companies employees resembles the regional distribution of legal companies (see Fig. 1) suggesting that legal companies among regions are quite similar to each other, so that there is no significant difference between simple count of companies and weighted count of companies with respect to the number of workers. The role of legal business in region's economy is distributed differently, though. It's clear that Moscow and Saint Petersburg, the two largest cities concentrating almost 1/3 of legal companies employees, are also characterized by the relatively highest portion of legal workers in total labor force. The majority of other regions is surprisingly uniform in having the small share of legal workers (0.1–0.2%) in labor force, with Sverdlovsk, Tyumen and Novosibirsk regions being rare exceptions. It is also worth noticing that regions with extremely low share in country's legal business (for example, Tyva republic and Altai territory) also usually have low share of legal workers in regional labor force, with a single exception of Evrei autonomous region. In addition, the regional role of legal business seems to have to connection with the presence or absence of multi-regional courts. However, looks like that the national role of a region in legal business is somehow connected with the location of such the courts, because almost all regions with relatively high share in country's legal companies employees (>1%) have at least one multi-regional court, and the reverse is true either: other regions typically do not have such the courts. Of course, there are regions that do not fall within this tendencies (Bashkortostan republic and Volgograd region, as well as three sub-Moscow regions, Vologda region and Zabaikal territory), but these deviations may be explained by various factors (e. g. for Tula, Vladimir and Kaluga regions the presence of multi-regional courts is not associated with the development of legal business probably due to the proximity to Moscow, so that the need in local legal firms does not arises).

<br>

```{r fig-3-employees-map, fig.cap="Fig 3. National and regional median yearly shares of legal companies employees in regions of Russia in 2018–2021"}
employees_map
```

<br>

The changes in the number of legal firms by region are depicted in Fig. 4. It can be clearly seen that the count of such companies increased in the majority of regions, indicating the development of legal business. Few regions have lost companies, all of them are located in the European part of Russia, and their net change in the number of legal firms from 2016 to 2021 is relatively low (usually less than 10%) compared to the increase rates in other regions. Sometimes neither decline or growth was not clearly expressed (for example, Altai republic, Lipetsk and Tambov regions). The regions with the largest growth are some republics at Notrh Caucas, as well as Kalmyk republic and Nenetz territory. Their results are probably explained by the low base effect, because the absolute counts of legal firms in all there regions are below 100 (see Fig. 1). As for other regions with high growth, Leningrad and Moscow regions should be remarked, because their results may be partially explained by the relocation of companies from the cities (Saint Petersburg and Moscow) to the surrounding regions. However, Moscow and Saint Petersburg also gained legal firms with one of the top net changes. The observed patterns of changes indicated by lines of the map can be divided into several groups. The first is the stable yet saturating increase typical for Moscow and Saint Petersburg regions, and, less expresses, Khabarovsk territory, Rostov and Ryazan regions. The second is the inverted U-shape curve which is charasteristic of the majority of regions with overall growth in the number of legal firms. This pattern can be explained by the coronavirus (COVID-19) pandemic effect that has negatively impacted small and medium-sized business and reverted the previous trend of the growth. The “height” of this inverted U varies by region (the highest observed in Chukotka territory, Tyva and Adygeya republics). It can be hypothesised that the first group is in fact a subset of the second, but the intrinsic growth rate before the pandemic was higher, resulting in the slowing down of the growth rather than in the final decrease. The third pattern is typical for the regions with overall decrease in the number of legal firms, and it can be described as “initial short-term growth, then continuous decline”. The fourth group is a mixture of increase and decrease, for example Nenetz territory and Northern Osetia republic.

<br>

```{r fig-4-timeline-map, fig.cap="Fig. 4. The change in the number of legal firms by region in 2016–2021"}
timeline_map
```

<br>

# Conclusion

The conducted regional analysis discovered several spatial features of legal business in Russia. First, the regions are quite different with respect to the count of legal firms or their employees. Several regional clusters with relatively high concentration of legal business can be found out, for example, the European South, Volga-Ural and Siberian. Unsurprisingly, Moscow and Saint Petersburg cities are characterized by the top count of legal firms and workers, but the overall concentration is really huge, because they jointly account for approximately 1/3 of Russia's legal businesses and legal companies employees. Besides, it's again Moscow and Saint Petersburg that have the highest share of legal companies employees in region's labor force (above 0.4% that is 2 or 3 times higher than the average value for other regions). However, there are some other regions with relatively high role of legal business in regional economy: Sverdlosvk, Tyumen and Novosibirsk.

Regions can be roughly grouped into two categories: one with the legal organisations prevailing in the legal business, and the other with individual entrepreneurs prevailing over companies. The high role of companies is typical for Moscow, Saint Petersburg, some Volga and Siberian regions, as well as a small portion of central regions (Vologda and Ivanovo), while individuals are important in the European South and North, in the majority of Central European regions, in several regions of the Urals, and in the Far East.

The role of region in country's legal business is usually associated with the presence of one or many multi-regional courts in the administrative center of this region. Only two regions do not act as a host for any of such the courts but still have relatively high level of development of legal business: Volgograd region and Bashkir republic. This observation may be explained in two ways: either the courts act as a magnets for legal business, because courts are necessary for many legal activities, or the courts are placed in regions and cities with highly developer legal business. Further research may throw the light on the direction of this court-business relation.

From 2016 to 2021, the number of legal firms in Russia's regions mostly increased. The decrease is characteristic only of few regions. The top growth was observed in Moscow, Saint Petersburg, as well as in the surrounding regions. In addition, some regions (mainly at the European South) also shown very high growth, but probably due to the low base effect. The yearly patterns of change may be categorized into four groups, and the inverted U-shape change explained likely by the coronavirus (covid-19) pandemic is the most common among the regions.

The results of this explanatory study provide an overview of Russia's recent legal business geography and may be used as a support for data-driven business and governmental decisions. Several research hypothesis also may be conducted from this analysis, thus this study may be treated as a base for further academic development in the field of legal services geography in Russia. Apart from this, the reproducible data-driven technique of this research based entirely on open data and open source software tools may be used (with minor changes) for regional analysis and mapping of other narrow service sector areas.

<br>

## References

<div id="refs"></div>